---
title: "A Report on Lake Ontario's Microbes"
author: "Mar Schmidt"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prepare the R Environment 

```{r packages}
# Load libraries/packages 
library(tidyverse)
```


# Load in the Lake Ontario Data 

```{r load-data}
# Load in lake ontario microbial community data 
sample_and_taxon <- 
  read_csv("data/sample_and_taxon.csv")

# Inspect the data 
glimpse(sample_and_taxon)
```


# Microbial Cell Abundance Versus Temperature 

```{r cells_vs_temp}
# temp on the X 
# cell abundance on the Y
# Colored by environmental Group
# make it look nice (e.g. labels, models, etc)

ggplot(data = sample_and_taxon, 
       aes(x = temperature, y = cells_per_ml)) + 
  geom_point(aes(color = env_group)) + 
  labs(x = "Temp (Celsius)", y = "Cell Density (cells/mL)") + 
  theme_classic() + 
  scale_color_manual(values = c("cornflowerblue", 
                                "goldenrod1",
                                "firebrick")) + 
  geom_smooth(method = "lm", color = "black") + 
  theme(legend.position = "bottom",
        legend.title = element_blank())
```

The above plot shows that: 

- There's a positive relationship between temperature and cell abundances. 
- For example, deep samples are the coldest and have the fewest cells. 

```{r data-summary, include = FALSE}
# How many samples do I have? 
n_samples <-
  sample_and_taxon %>%
  # Only pick out sample ID column
  select(sample_id) %>%
  # Number of distinct names in the column 
  n_distinct()

# Calculate the minimum and maximum temps
# min
minTemp <- 
  sample_and_taxon %>%
  summarize(round(min(temperature), digits = 1)) %>%
  pull()

# max
maxTemp <- 
  sample_and_taxon %>%
  summarize(round(max(temperature), digits = 1)) %>%
  pull()
```

The total number of samples is `r n_samples`. For this set of samples, temperature ranged from a minimum of `r minTemp` celsius to a maximum of `r maxTemp` celsius.



